OCMu64: a solver for One-sided Crossing Minimization

### 2024-06-19 tex/description.tex
add DOI to description
Changes in description intro

\author{Mees de Vries}{}{}{}{}

@@

\relatedversion{} %optional, e.g. full version hosted on arXiv, HAL, or other respository/website
%\relatedversiondetails[linktext={opt. text shown instead of the URL}, cite=DBLP:books/mk/GrayR93]{Classification (e.g. Full Version, Extended Version, Previous Version}{URL to related version} %linktext and cite are optional

@@

\newcommand{\R}{\mathbb R}
\renewcommand{\b}{\prec}
\newcommand{\be}{\preceq}
\newcommand{\g}{\sim}
\renewcommand{\u}{\overline{u}}
\renewcommand{\v}{\overline{v}}
\newcommand{\w}{\overline{w}}

@@

The 2024 edition of PACE, an annual optimization challenge, considers the
\emph{one-sided crossing minimization} problem, defined as follows.
Given is a bipartite graph $(A, B)$ that is drawn in the plane at points
$(0, i)$ and $(1,j)$ for components $A$ and $B$ respectively. The order of $A$
is fixed, and the goal is the find an order of $B$ that minimizes the number of
crossings when edges are drawn as straight lines.

We introduce some new reductions and overview our algorithm. Proofs are brief or
omitted due to lack of space.

\section{Definitions}
Since the order of $A$ is fixed, we set $A=\{0, \dots, [A]\}$ (where $[n]=\{0,\dots,n-1\}$) and draw the
vertices in this order.
Let $u, v\in B$ be two vertices in $B$. Then $N(u),N(v)\subseteq A$ are their sets of
neighbours.
We write $c(u,v)=\sum_{a\in N(u)}\sum_{b\in N(v)} [a>b]$ for the number of crossings between edges from $u$ and $v$
when $u$ is drawn before $v$. For $X,Y\subseteq B$ we set
$c(X,Y) = \sum_{x\in X}\sum_{y\in Y} c(x,y)$ for the cost of ordering all
vertices of $X$ before all vertices of $Y$.
More generally, $c(X,Y,Z) = c(X, Y) + c(X, Z) + c(Y, Z)$.
We also consider the \emph{reduced cost} $r(X,Y) = c(X, Y) - c(Y, X)$, which is
negative when $X$ is better \emph{before} $Y$ and positive when $X$ is better \emph{after} $Y$.

We write $u\b v$ when $u$ must come before $v$ in all minimal solutions and
$u\be v$ when there exist minimal solutions where $u$ comes before $v$, and we
call such $(u,v)$ \emph{dominating pairs}.
We write $u\g v$ when there is a minimal solution where $u$ and $v$ are
consecutive, in that order.

@@

% TODO: Citation
\begin{observation}\label{lem:weak}
  When all neighbours of $u$ come before all neighbours of $v$, then $u$ goes
  before $v$. I.e., when $\max N(u) \leq \min N(v)$, then $u\be v$.
  When additionally ${\min N(u) < \max N(v)}$, then $u\b v$.
\end{observation}

>>>>

\author{Mees de Vries}{Unaffiliated, Netherlands}{meesdevries@protonmail.com}{}{}

@@

\relatedversion{\url{https://doi.org/10.5281/zenodo.11671980}} %optional, e.g. full version hosted on arXiv, HAL, or other respository/website
%\relatedversiondetails[linktext={opt. text shown instead of the URL}, cite=DBLP:books/mk/GrayR93]{Classification (e.g. Full Version, Extended Version, Previous Version}{URL to related version} %linktext and cite are optional

@@

\newcommand{\R}{\mathbb R}
\renewcommand{\b}{\prec}
\newcommand{\be}{\preceq}
\newcommand{\g}{\sim}
\renewcommand{\u}{\overline{u}}
\renewcommand{\v}{\overline{v}}
\newcommand{\w}{\overline{w}}
\DeclareMathOperator{\supp}{supp}

@@

The 2024 edition of PACE, an annual optimization challenge, considers the
\emph{one-sided crossing minimization} problem, defined as follows.
Given is a bipartite graph $(A, B)$ that is drawn in the plane at points
$(i, 1)$ and $(j,0)$ for components $A$ and $B$ respectively. The ordering of $A$
is fixed, and the goal is the find an ordering of $B$ that minimizes the number of
crossings when edges are drawn as straight lines.

We introduce some new reductions and give an overview of our algorithm. Proofs are brief or
omitted due to lack of space.

\section{Definitions}
We use $<$ to compare vertices in $A$ in their fixed ordering. We generalize to weighted graphs
for the proof of \ref{lem:sdp}: a node $u \in B$ is taken to be a function $u: A \to \mathbb
R^{\geq 0}$, where weight 0 is an absent edge. We write $c(u, v) = \sum_{(a, b) \in A^2}
u(a)v(b)[a > b]$; in the unweighted case, this is the \emph{crossing number}, the number of
crossings between edges incident to $u$ and $v$ when $u$ is drawn before $v$.

For $X,Y\subseteq B$ we set $c(X,Y) = \sum_{x\in X}\sum_{y\in Y} c(x,y)$ for the cost of
ordering all vertices of $X$ before all vertices of $Y$. More generally, $c(X,Y,Z) = c(X, Y) +
c(X, Z) + c(Y, Z)$. We also consider the \emph{reduced cost} $r(X,Y) = c(X, Y) - c(Y, X)$,
which is negative when $X$ is better \emph{before} $Y$ and positive when $X$ is better
\emph{after} $Y$.

We write $u\b v$ when $u$ must come before $v$ in all minimal solutions, and say that $(u, v)$
is a \emph{fixed} pair. When there exists a minimal solution where $u$ comes before $v$, we
write $u \be v$ and call $(u, v)$ \emph{weakly fixed}.
We write $u\g v$ when there exists a minimal solution where $u$ and $v$ are
consecutive, in that order.

@@

% TODO: Citation
\begin{observation}\label{lem:weak}
  When all neighbours of $u$ come before all neighbours of $v$, then $u$ goes before $v$. That
    is, when $\max \supp (u) \leq \min \supp (v)$, then $u\be v$, and in fact $u \b v$ holds
    unless $u$ and $v$ are both connected to only one and the same node in $A$.
\end{observation}
